### 2021-08-11 Introduction.tex

\section{Introduction}
\label{introduction}

Covariance matrix estimation is a fundamental statistical problem that plays an essential role in various applications, such as portfolio management \citep{ledoit2003improved}, genomics \citep{schafer2005shrinkage}, and array signal processing \citep{abramovich2001locally}. However, in modern problems the number of features can be of the same order as or exceed the sample size, and the standard sample covariance matrix estimator behaves poorly in this regime. To overcome these issues, various methods have been developed to estimate high-dimensional covariance matrix. These can roughly be divided into two groups, according to whether they impose assumptions about the structure of population covariance matrix.

Structured methods make structural assumptions about the population covariance matrix. One class models the population covariance matrix as sparse. The most common method to address this problem is thresholding \citep{rothman2009generalized, cai2011adaptive}. Penalized likelihood methods \citep{xue2012positive} can also estimate large-scale sparse covariance matrix by penalizing a log-likelihood function. Another class of methods assume the data arise from a factor model \citep{fan2008high}, so that the covariance matrix has low intrinsic dimension. Other common structured methods assume that the covariance matrix is banding \citep{li2017estimation} or Toeplitz matrix \citep{liu2017covariance}.

In contrast, unstructured methods do not make any assumptions about the population covariance matrix, yet can still outperform the sample covariance matrix. A first example was the linear shrinkage approach of \citet{ledoit2004well}, which shrinks the sample covariance matrix toward a scaled identity matrix. More recently, nonlinear shrinkage methods were developed \citep{ledoit2012nonlinear, ledoit2019quadratic, lam2016nonparametric}. These shrink the eigenvalues of the sample covariance matrix toward clusters. Linear shrinkage can be viewed as a special case of nonlinear shrinkage, as it shrinks sample eigenvalues toward their global mean.

Nonlinear shrinkage estimators have desirable optimality properties \citep{ledoit2018analytical} and show excellent performance. However, they modify only the sample eigenvalues and not the sample eigenvectors. It is known that sample eigenvectors are not consistent estimators of population eigenvectors when the dimension and the sample size increase at the same rate \citep{mestre2008asymptotic}. This suggests that there may exist a class of unstructured estimators that can outperform nonlinear shrinkage.

%In this paper, we propose a new covariance matrix estimator which is approximately optimal among all separable estimators. Our method is motivated by the similarity between covariance estimation and vector mean estimation. Covariance matrix estimation under Frobenius loss can be viewed as a vector estimation problem by converting the matrix to a vector. \citep{cai_liu} used this motivation to generalize thresholding method in mean vector estimation to large sparse covariance matrix estimation. Nevertheless, they are not totally the same because covariance matrix has repeated entries due to its symmetry. In mean vector estimation problem, optimality has been achieved among all separable decision rules. \citep{jiang_zhang2009} and \citep{brown_greenshtein} proposed empirical Bayes methods to approximate the optimal separable estimator. We proposed to apply vector estimation methods to covariance estimation problem and expect that these methods will also have good properties here. Result in simulation and real data set shows that our estimator has good performance in different scenarios.

Here we propose a new unstructured estimator for high-dimensional covariance matrices. Our approach centers on vectorizing the covariance matrix and treating matrix estimation as a vector estimation problem. We do this because it allows us to use a nonparametric empirical Bayes shrinkage procedure, which has been shown in the compound decision literature to have excellent properties \citep{jiang2009general, koenker2014convex}. We then reassemble the estimated vector into matrix form and project onto the space of positive-definite matrices to give our final estimator. Surprisingly, though our vectorized approach essentially ignores the matrix structure, it can still substantially outperform a number of state-of-the-art proposals in simulations and a real data analysis.

The article is organized as follows. In Section \ref{sec:method}, we briefly review compound decision theory and then introduce our proposed approach. In Section \ref{numerical results} we illustrate the performance of our method in simulations and a gene expression dataset. Finally, Section \ref{sec:discussion} concludes with a discussion. Our procedure is implemented in the R package \verb|cole|, available on GitHub.